---
title: "Querying the Data"
---

<br>

For all examples the movies data set contained in the package will be used.
```{r}
library(UpSetR)
movies <- read.csv( system.file("extdata", "movies.csv", package = "UpSetR"), header=T, sep=";" )
```

<br>

*queries* Parameter Breakdown
----------------------------------
Each list contained in the **queries** parameter takes 4 fields: query, params, color, and active.

- *query* specifies which query is going to be run

- *params* is a list of paramters for the query to work on

- *color* is the color that will represent the query on the plot. If no color is provided one will be selected from a color palette.

- *active* determines how the query will be represented on the plot. If active is TRUE, the intersection size bar will be overlayed by a bar representing the query. If active is FALSE, a jitter point will be placed on the intersection size bar.

To see how queries can be explored and visualized on an element level visit the **Attribute Plots** vignette.

<br>

Example 1: Built-in Intersection Query
-------------
This example shows how to use the built in intersection query, **intersects**, to find or display elements in specific intersections.
```{r, fig.width=12}
upset(movies, queries = list(list(query = intersects, params = list("Drama", "Comedy", "Action"), color = "orange", active = T), list(query = intersects, params = list("Drama"), color = "red", active = F), list(query = intersects, params = list("Action", "Drama"), active = T)))
```

<br>

Example 2: Built-In Elements Query
-------------
This example shows how to use the built in element query, **elements**, to visualize how certain elements are distributed amongst the intersections.
```{r, fig.width=12}
upset(movies, queries = list(list(query = elements, params = list("AvgRating", 3.5, 4.1), color = "blue", active = T), list(query = elements, params = list("ReleaseDate", 1980, 1990, 2000), color = "red", active = F)))
```

<br>

Example 3: Using Expression Parameter to Subset Intersection and Element Queries
-------------
This example shows how to use the **expression** parameter to subset element and intersection queries.
```{r, fig.width=12}
upset(movies, queries = list(list(query = intersects, params = list("Action", "Drama"), active = T), list(query = elements, params = list("ReleaseDate", 1980, 1990, 2000), color = "red", active = F)), expression = "AvgRating > 3 & Watches > 100")
```

<br>

Example 4: Creating Custom Queries on Set Elements and Attributes
-------------
Creating a custom query to operate on the rows of the data.
```{r}
Myfunc <- function(row, release, rating){
    data <- (row["ReleaseDate"] %in% release) & (row["AvgRating"] > rating)
}
```

Applying the created query to the queries parameter.
```{r, fig.width=12}
upset(movies, queries = list(list(query = Myfunc, params = list(c(1970,1980, 1990, 1999, 2000), 2.5), color = "blue", active =T)))
```

<br>

Example 5: Applying Everything at Once
--------------
Combining pieces from all previous examples into one awesome query!
```{r, fig.width=12}
upset(movies, queries = list(list(query = Myfunc, params = list(c(1970,1980, 1990, 1999, 2000), 2.5), color = "orange", active =T), list(query = intersects, params = list("Action", "Drama"), active = F), list(query = elements, params = list("ReleaseDate", 1980, 1990, 2000), color = "red", active = F)), expression = "AvgRating > 3 & Watches > 100")
```

